---
title: Regression Analysis on the Relationship Between Advertising Budgets and Product
  Sales
author: "Shannon Chang"
date: "October 14, 2016"
output: pdf_document
---
```{r, eval = TRUE, echo = FALSE, message = FALSE}
library(xtable)
library(png)
library(grid)
library(gridExtra)
load("../data/regression.RData")
load("../data/correlation-matrix.RData")
```



# Abstract
In this report, I will reproduce the multiple linear regression analysis detailed 
in Chapter 3 of Gareth James, Daniela Witten, Trevor Hastie, and Robert Tibshirani's 
[_An Introduction to Statistical Learning_](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20First%20Printing.pdf). 
The analysis is carried out on the 
[`Advertising.csv`](http://www-bcf.usc.edu/~gareth/ISL/Advertising.csv) dataset 
that is paired with the textbook, and contains data on product sales in over two 
hundred different markets along with the advertising budgets for the product in 
each market by different mediums: `TV`, `Radio`, and `Newspaper`.  
  
Specifically, I will reproduce:  
* __Table 3.1__ (page 72) - Coefficient estimates of simple linear regression 
models of `Sales` on `TV`, `Sales` on `Radio`, and `Sales` on `Newspaper`  
* __Table 3.4__ (page 74) - Least squares coefficient estimates of the multiple 
linear regression model of `Sales` on `TV`, `Radio`, and `Newspaper`  
* __Table 3.5__ (page 75) - A correlation matrix for `TV`, `Radio`, `Newspaper`, and 
`Sales`  
* __Table 3.6__ (page 76) - $RSE$, $R^2$, and $F-statistic$ values from the least 
squares model for the regression of `Sales` on `TV`, `Radio`, and `Newspaper`





# Introduction
Suppose a company wants advice on how to increase sales for one of its products. 
There is, of course, no concrete way to insure increased sales, but we can 
influence greater sales through advertising. Imagine that we are statistical 
consultants hired for this project. To convince the company to invest in 
advertising campaigns, we must first prove to the company that there is a 
relationship between advertising budget and sales. From here, we can then advise 
the company on appropriate budgets to better reach sales targets. Thus, 
the goal for this analysis is to determine whether there is a relationship 
between advertising budget and sales and, if so, how strong the relationship is 
between the two. Given that there is a relationship between advertising budget and 
sales, we would want to construct an accurate model that can be utilized to predict 
sales based on the size of advertising budget. There are three mediums detailed in 
`Advertising.csv`: `TV`, `Radio`, and `Newspaper`, so we will examine their individual 
relationships with `Sales` and then their combined relationship with `Sales`. 





# Data
The `Advertising.csv` dataset consists of advertising budgets (in thousands of 
dollars) by medium: `TV`, `Radio`, and `Newspaper`. Product sales (in thousands
of units) are listed under `Sales`. There are 200 rows of data, indicating 200 
different markets. A very general overview of the distribution for each column of data 
can be seen in the histograms below:   
Histograms for the two columns are shown below:  
```{r, eval = TRUE, echo = FALSE, message = FALSE}
tv <- rasterGrob(as.raster(readPNG("../images/histogram-tv.png")), interpolate = FALSE)
radio <- rasterGrob(as.raster(readPNG("../images/histogram-radio.png")), interpolate = FALSE)

grid.arrange(tv, 
             radio, 
             ncol = 2)

np <- rasterGrob(as.raster(readPNG("../images/histogram-newspaper.png")), interpolate = FALSE)
sales <- rasterGrob(as.raster(readPNG("../images/histogram-sales.png")), interpolate = FALSE)

grid.arrange(np, 
             sales, 
             ncol = 2)
```





# Methodology
## Setting Up a Model
To examine the association between `Sales` and `TV`, `Radio`, and `Newspaper`, whether 
individually or altogether, we model the relationship by _linear regression_. For the 
association with `Sales` and the advertising budget for each medium, we would model the 
relationship by _simple linear regression_. For the association between `Sales` and 
all three mediums, we would model the relationship by _multiple linear regression_. 

### Simple Linear Regression 
This method involves predicting a quantitative response $Y$, based on the predictor 
variable $X$. For the association between `Sales` and the advertising budget for each 
medium, we would model the relationships as:  
$$Sales = \beta_0 + \beta_1TV$$  
$$Sales = \beta_0 + \beta_1Radio$$  
$$Sales = \beta_0 + \beta_1Newspaper$$  
Here, Here, $\beta_0$ represents the intercept of the linear model while $\beta_1$ 
represents the slope.  
  
Since all $\beta_0$ and $\beta_1$ are unknown for each association, we would need to 
calculate estimates for the their coefficients instead. In a visual sense, we would 
want to graph all the data for `TV` and `Sales` and fit a line 
$Sales = \beta_0 + \beta_1TV$ as close as possible to our 200 data points. We would 
repeat this process for `Radio` and `Newspaper` as well. Then, we can optimize the 
fit of each line using the _least squares criterion_. This involves minimizing the _residual sum of squares_; a _residual_ is the distance between each data point and its predicted value from the linear model). The linear model/line that we fit would be based on an average of the squares.  
  
From a computational perspective, we can start fitting the line by calculating
estimates $\hat{\beta_0}$ and $\hat{\beta_1}$ for $\beta_0$ and $\beta_1$. 
$\hat{\beta_0}$ and $\hat{\beta_1}$ are chosen to minimize the aforementioned 
_residual sum of squares_ (_RSS_), which is given by the equation: 
$$RSS = (y_1 - \hat{\beta_0} - \hat{\beta_1}x_1)^2 + (y_2 - \hat{\beta_0} - \hat{\beta_1}x_2)^2 + ... + (y_n - \hat{\beta_0} - \hat{\beta_1}x_n)^2$$
$\hat{\beta_0}$ and $\hat{\beta_1}$, in turn, are shown through caculus derivation 
as: 
$$\hat{\beta_0} = \bar{y} - \hat{\beta_1}\bar{x}$$
$$\hat{\beta_1} = \frac{\sum(x_i-\bar{x})(y_i - \bar{y})}{\sum(x_i-\bar{x})^2}$$ 
Here, $\bar{x}$ is the mean of `TV` (or `Radio` or `Newspaper`), while $\bar{y}$ 
is the mean of `Sales`. 

To evaluate the accuracy of these estimates, we start by calculating standard errors 
of the standard means. We can then use these standard errors to perform 
_hypothesis tests_ on the estimates. Thus, we would be testing the _null hypothesis_ 
that $$H_0: There\: is\: no\: relationship\: between\: TV\:  and\:  Sales$$ versus 
the _alternative hypothesis_ that 
$$H_1: There\: is\: some\: relationship\: between\: TV\: and\: Sales$$   
Here, _TV_ can be substituted with _Radio_ and _Newspaper_.  
  
Numerically, we would be testing $$H_0: \beta_1 = 0$$ versus 
$$H_0: \beta_1 \neq 0$$  
To do so, we would calculate a _t-statistic_:
$$t = \frac{\hat{\beta_1} - 0}{SE(\hat{\beta_1})}$$ 
This measures the number of standard deviations that our estimate for $\beta_1$ is 
away from 0.  
  
From this, we can calculate a _p-value_, which is the probability of observing 
any value greater than or equal to |_t_|. A small p-value would indicate that it 
is unlikely to observe a meaningful association between the predictor (`TV`, or 
`Radio`, or `Newspaper`) and the response (`Sales`) purely by chance without some 
true relationship between the two. Thus, a small p-value would allow us to 
_reject the null hypothesis_ and determine that there is a relationship between 
`TV` (or `Radio` or `Newspaper`) and `Sales`. In general, 5% or 1% are used 
as p-value benchmarks. 

### Muliple Linear Regression
This method is, in many ways, similar to __simple linear regression__, but with the 
added complexity of more variables. We start by predicting a quantitative response 
$Y$, based on the predictor variables $X_1$, $X_2$, and $X_3$. For the combined 
association between `Sales` and `TV`, `Radio`, and `Newspaper`, we would model this 
relationship as: 
$$Sales = \beta_0 + \beta_1TV + \beta_2Radio + \beta_3Newspaper$$  
Here, Here, $\beta_0$ represents the intercept of the linear model while $\beta_1$, 
$\beta_2$, and $\beta_3$ represent the three different slopes. Each $\beta_j$ can 
be interpreted as 
*the average effect on $Y$ of a one unit increase in $X_j$, holding all other predictors fixed*.Note that fitting three simple linear regressions is not the same as fitting one 
multiple linear regression for `Sales`, because the each simple linear regression would 
not take into account the effects of the other two mediums and, subsequently, any 
interaction effects.  
  
Visualizing this model is a complicated matter, as each additional predictor adds 
another dimension to the graph. But much like in the _simple linear regression_, we 
are once again trying the minimize the sum of squared residuals, this this time 
with an expanded formulation: 
$$
\begin{aligned}
  RSS &= \sum_{i=1}^{n} (y_i - \hat{y_i})^2 \\
  &= \sum_{i=1}^{n} (y_i - \hat{\beta_0} - \hat{\beta_1}x_{i1} - \hat{\beta_p}x_{i3}) 
\end{aligned}
$$ 
The coefficient estimates of $\hat{\beta_0}$, $\hat{\beta_1}$, $\hat{\beta_2}$, and 
$\hat{\beta_3}$ are derived more complicatedly through matrix algebra and will not 
be listed here. 
  
Hypothesis testing for similar to that conducted for _simple linear regression_ can 
also be conducted by _multiple linear regression_. The calculated results, however, 
will change because of the way in which $\hat{\beta_0}$, $\hat{\beta_1}$, 
$\hat{\beta_2}$, and $\hat{\beta_3}$ are calculated. Hence, results of a hypothesis 
test conducted from the _multiple linear regression_, say for `Sales` and `TV` could 
have different results compared to a hypothesis conducted from a 
_simple linear regression_ for the same variables.  
  
In general for _multiple linear regression_, however, we want to test whether all of 
the regression coefficients are zero, i.e. whether $\beta_1 = \beta_2 = \beta_3 = 0$ 
in this case. Hence, we would test the null hypothesis that 
$$H_0: \beta_1 = \beta_2 = \beta_3 = 0$$ 
versus the _alternative hypothesis_ that 
$$H_1: At\: least\: one\: \beta_j\: is\: non-zero.$$ 
This hypothesis would be conducted by calculating the $F-statistic$: 
$$F = \frac{(TSS - RSS)/p}{RSS/(n-p-1)}$$
(TSS is further explained in the sections below.) 
  
We can also test the hypothesis that a certain subset $q$ of the coefficients are 0. 
For this, we would test the null hypothesis that
$$H_0: \beta_{p-q+1} = \beta_{p-q+2} = \beta_{p} = 0$$
where p is the number of predictors, which in this case is 3. 
The $F-statistic$ would then be calculated by: 
$$F = \frac{(RSS_0 - RSS)/q}{RSS/(n-p-1)}$$ 
where $RSS_0$ is the residual sum of squares for a second model that uses all the 
variables _except_those last $q$. 
  
## Evaluating Accuracy of the Model
After conducting a hypothesis test, we will want to examine the extent to which 
the model fits the data. There are three values that we can look at to assess this: 
the _residual standard error_ ($RSE$), the $R^2$ value, and the $F-statistic$. 

### Residual standard error (RSE)
The RSE is an estimate of the standard deviation of errors, the distances from 
each data point to its predicted value based on the linear model we fit. In other 
words, it is the average amount that the response (`Sales`) will differ from 
the true regression line and is given by the formula: 
$$RSE = \sqrt{\frac{1}{n-2}RSS} = \sqrt{\frac{1}{n-2}\sum{(y_i - \hat{y_i})^2}}$$

### $R^2$ value
The $R^2$ statistic is, technically speaking, the proportion of variance explained 
by our fitted model. Specifically, it measures the _proportion of variability in the response (`Sales`) that can be explained using the predictor (`TV`)_. 
The closer $R^2$ is to 1, the greater the proportion of variability that is 
explained. Its formula is given by: 

$$ R^2 = \frac{(TSS - RSS)}{TSS} = 1 - \frac{RSS}{TSS}$$  
Here, the _total sum of squares_, TSS = $\sum(y_i -\bar{y})^2)$ measures the total 
variance in the response $Y$, and can be thought of as the amount of variability 
that already exists in the response, even before we perform any regression analysis. 
Thus, the R^2 value is a ratio of variability in $Y$ that can be explained by our 
model to the variability that exists inherently in $Y$. 

### $F$-statistic
Following up on the earlier discussion of the $F-statistic$, it can be shown by linear 
model assumptions that 
$$E{RSS/(n-p-1)} = \sigma^2$$
Provided that $H_0$ is true, 
$$E{(TSS-RSS)/p} = \sigma^2$$
Thus, when there is no relationship between the response and its predictors, we can 
expect the $F-statistic$ to be close to or equal to 1. If $H_1$ is true, on the other 
hand, then $E{(TSS-RSS)/p} > \sigma^2$, and we can expect F to be greater than 1. 





## Results
### Simple Linear Regression
For the association between `Sales` and `TV`, I used R to calculate a regression 
object, which produced the coefficient estimates for the model and necessary 
calculations for performing a _t-test_. I also plotted the observed data against the line fitted by the regression object in order to obtain a visualization of this 
analysis. The code is located in `code/scripts/regression-script.R" of the 
repository for this paper, and is summarized by the following: 

```{r, eval = FALSE}
# Load necessary package(s) and data
library(readr)
advertising <- read.csv(file = "../data/Advertising.csv", row.names = 1)

# Generate regression object
sales_tv_reg <- lm(Sales ~ TV, data = advertising)

# Generate summary
sales_tv_sum <- summary(sales_tv_reg)

# Generate plot
plot(x = advertising$TV, y = advertising$Sales, 
     xlab = "TV Advertising Budgets (in thousands of dollars)", 
     ylab = "Sales (in thousands of product units)")
abline(sales_tv_reg)
```

This generated the following outputs: 
```{r, eval = TRUE, echo = FALSE, results = "asis", message = FALSE}
sales_tv <- xtable(sales_tv_sum$coefficients, 
                      digits = c(0, 3, 3, 2, 4), 
                      caption = c("Coefficient Estimates from the Simple Regression of Sales on TV"))
print.xtable(sales_tv, comment = FALSE, caption.placement = "top")
```

```{r, eval = TRUE, echo = FALSE, fig.width = 4, fig.height = 4, fig.align='center'}
tv_sales <- readPNG("../images/scatterplot-tv-sales.png")
grid.raster(tv_sales)
```



```{r, eval = TRUE, echo = FALSE, fig.width = 4, fig.height = 4, fig.align='center'}
radio_sales <- readPNG("../images/scatterplot-radio-sales.png")
grid.raster(radio_sales)
```

```{r, eval = TRUE, echo = FALSE, fig.width = 4, fig.height = 4, fig.align='center'}
np_sales <- readPNG("../images/scatterplot-newspaper-sales.png")
grid.raster(np_sales)
```

```{r, eval = TRUE, echo = FALSE, fig.width = 6, fig.height = 6, fig.align='center'}
matrix <- readPNG("../images/scatterplot-matrix.png")
grid.raster(matrix)
```

```{r, eval = TRUE, echo = FALSE, fig.width = 4, fig.height = 4, fig.align='center'}
residual <- readPNG("../images/residual-plot.png")
grid.raster(residual)
```

```{r, eval = TRUE, echo = FALSE, fig.width = 4, fig.height = 4, fig.align='center'}
scale_location <- readPNG("../images/scale-location-plot.png")
grid.raster(scale_location)
```

```{r, eval = TRUE, echo = FALSE, fig.width = 4, fig.height = 4, fig.align='center'}
normal_qq <- readPNG("../images/normal-qq-plot.png")
grid.raster(normal_qq)
```




```{r, eval = TRUE, echo = FALSE, results = "asis", message = FALSE}
sales_radio <- xtable(sales_radio_sum$coefficients, 
                      digits = c(0, 3, 3, 2, 4), 
                      caption = c("Coefficient Estimates from the Simple Regression of Sales on Radio"))
print.xtable(sales_radio, comment = FALSE, caption.placement = "top")
```

```{r, eval = TRUE, echo = FALSE, results = "asis", message = FALSE}
sales_newspaper <- xtable(sales_np_sum$coefficients, 
                      digits = c(0, 3, 3, 2, 4), 
                      caption = c("Coefficient Estimates from the Simple Regression of Sales on Newspaper"))
print.xtable(sales_newspaper, comment = FALSE, caption.placement = "top")
```

```{r, eval = TRUE, echo = FALSE, results = "asis", message = FALSE}
multi <- xtable(multi_sum$coefficients, 
                      digits = c(0, 3, 3, 2, 4), 
                      caption = c("Least Squares Coefficient Estimates from the Multiple Linear Regression of Sales on TV, Radio, and Newspaper"))
print.xtable(sales_radio, comment = FALSE, caption.placement = "top")
```

```{r, eval = TRUE, echo = FALSE, results = "asis", message = FALSE}
cor_table <- xtable(adv_cor, 
                    caption = "Correlation Matrix")
print.xtable(cor_table, comment = FALSE, caption.placement = "top")
```

## Conclusions
